---
title: "Open Policing Project in Witchita Kansas"
author: Kyle Helie
output:
  html_document:
    code_folding: hide
---

## Setup

First, let's load the necessary libraries and data that will allow us to
begin our investigation!

```{r setup, message=FALSE,  warning=FALSE}
## Libraries to include
library(tidyverse)
library(lubridate)

## Load the data
# Replace the path below with the path to where your data lives
data_path <- "https://datajournalism.tech/wp-content/uploads/2019/10/wichita.csv"
stops <- read_csv(data_path)

# Additional data and fixed values we'll be using
population_2016 <- tibble(subject_race = c("asian/pacific islander", "black", "hispanic", "other/unknown","white"),
  num_people = c(19294, 42485, 65090, 16686, 245499)) %>% 
  mutate(subject_race = as.factor(subject_race))

center_lat <- 37.689820
center_lng <- -97.336454
```

## Exploratory data analysis


```{r}
colnames(stops)
nrow(stops)
str(stops)
summary(stops)
```




```{r}

# This method uses the group_by/summarize paradigm
stops %>% 
  group_by(subject_race) %>% 
  summarize(
    n = n(),
    prop = n / nrow(.)
  )
```


``

## Benchmark test

We saw before that over two-thirds of stops were of black drivers. The by-race
stop counts are only meaningful, though, when compared to some baseline. If
the Philadelphia population was about two-thirds black, then two-thirds of stops
being of black drivers wouldn't be at all surprising. 

### Stop rates

In order to do this baseline comparison, we need to understand the racial
demographics in our Philly population data. The data as we've given it to you
has raw population numbers. To make it useful, we'll need to compute the
_proportion_ of Philadelphia residents in each demographic group. (Hint: use the
`mutate()` function.)

```{r}
population_2016 %>% 
  mutate(prop = num_people / sum(num_people))
```



```{r}
stop_final <- stops %>% 
  count(subject_race) %>% 
  left_join(
    population_2016,
    by = "subject_race"
  ) %>% 
  mutate(stop_rate = n / num_people)
```
## Data visualizations

```{r}

bar <- ggplot(stop_final,
       aes(x=reorder(subject_race,stop_rate), y=stop_rate))+
  geom_bar(stat="identity", 
           position="identity", 
           fill="red")+
  geom_hline(yintercept = 0) +
  labs(title="Stopped Drivers by Race",
       subtitle = "African American drivers were stopped more than White people in the city of Wichita,Kansas")+
  coord_flip()

options(scipen=10000)

bar
```
```{r}
library(leaflet)
library(httpuv)

#Step 1. Create a color palette of your choice.
race <- colorFactor(c("beige", "black", "orange", "darkgreen", "blue"),
domain=c("white", "black", "asian/pacific islander", "hispanic", "other/unknown"),
ordered = TRUE)


#Step 3. Drop missing data values.
f <- stops %>% drop_na(lat, lng)

#Step 4. Map the data set.
leaflet(f) %>%
  addProviderTiles(providers$CartoDB) %>% 
  setView(lng= -97.336454, lat= 37.689820, zoom=16) %>% 
  addCircleMarkers(~lng,
                   ~lat,
                   popup=paste("This is a/an", f$subject_race, "and", f$subject_sex, "driver."),
                   weight= 3, 
                   radius=4, 
                   color=~race(subject_race),
                   stroke=F,
                   fillOpacity = 1)
```

